Real Estate

### 2022-11-01 src/demo_data_wrangling.Rmd
added median rent data and weights for data

>>>>

2021 Median Rent from https://www.huduser.gov/portal/datasets/50per.html#2021

```{r}
# median rents from hud
MSA_median_rent2021_data <- read_csv(here("data", "raw", "2021MSA_medianrent.csv")) %>% 
  separate(col = cbsasub21,
           into = c("metro", "cbsa"),
           sep = 5) %>% 
  separate(col = cbsa,
           into = c("cbsa", "unknown"),
           sep = -5) %>% 
  mutate(cbsa = as.double(cbsa)) %>% 
  select(cbsa, rent50_0, rent50_1, rent50_2, rent50_3, rent50_4) %>% 
  right_join(potential_MSAs) 
```

### 2022-11-01 src/real_estate.Rmd
added median rent data and weights for data

city_realEstate_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990")
```

```{r}
# select columns in demographic data
MSA_demo_data <- MSA_demo_data %>% 
  clean_names() %>% 
    # calculate income to home price
  mutate(income_to_homeprice = median_income_usd_2021/house_price_2021) %>% 
  select(cbsa, msa, state, employment_yo_y21_22, pop_percent_change20_21, median_income_usd_2021, income_to_homeprice)
```

@@

```{r}
# combine real estate and demographic data by cbsa
MSA_realEstate_data <- MSA_demo_data %>% 
  left_join(MSA_realEstate_data) %>% 
  # calculate rent to income ratio
  mutate(rent_to_income = avg_rent_2021/median_income_usd_2021)

@@

```{r}
```

```{r}
# calculate income to home price
```

```{r}
# normalize columns into new data table
n_MSA_realEstate_data <- MSA_realEstate_data %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>% 
  # normalize employment growth
  mutate(n_employ_change = (employment_yo_y21_22 - min(employment_yo_y21_22)) / (max(employment_yo_y21_22) - min(employment_yo_y21_22))) %>% 
  # normalize rent to income
  mutate(n_rent_to_income = (rent_to_income- min(rent_to_income)) / (max(rent_to_income) - min(rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice)))

@@

```{r}
# invert normalized rent to income ratio so lower values (more favorable) have higher scores

= 1- normalized score
```

```{r}
# weights from owen

# place holder column
MSA_real_estate_score <- 
  mutate

weights <- c()

@@

>>>>

city_realEstate_data <- read_sheet("https://docs.google.com/spreadsheets/d/10n55Gt0rySN96OGTQ2_8PzeTcX66n26sOlZBiuiAeBU/edit#gid=808954990")

MSA_median_rent_data <- read_sheet("https://docs.google.com/spreadsheets/d/1oxNxWQRWpANbgb603ShBlw84QkyvsRaH4TzHKKVWhxw/edit#gid=0")
```

```{r}
# select columns in demographic data
MSA_demographics <- MSA_demo_data %>% 
  clean_names() %>% 
    # calculate income to home price
  mutate(income_to_homeprice = median_income_usd_2021/house_price_2021) %>% 
  select(cbsa, msa, state, employment_yo_y21_22, pop_percent_change20_21, median_income_usd_2021, income_to_homeprice)
```

@@

```{r}
# MSA_median rent data
MSA_realEstate_data<- MSA_median_rent_data %>% 
  select(cbsa, rent50_1) %>%
  mutate(rent50_1 = as.numeric(rent50_1)) %>% 
  left_join(MSA_realEstate_data)
```

```{r}
# combine real estate and demographic data by cbsa
MSA_realEstate_data <- MSA_demographics %>% 
  left_join(MSA_realEstate_data) %>% 
  # calculate rent to income ratio
  mutate(avg_rent_to_income = avg_rent_2021/median_income_usd_2021) %>% 
  mutate(med_rent_to_income = rent50_1/median_income_usd_2021)

@@

```{r}
# normalize columns into new data table
MSA_realEstate_data <- MSA_realEstate_data %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>% 
  # normalize employment growth
  mutate(n_employ_change = (employment_yo_y21_22 - min(employment_yo_y21_22)) / (max(employment_yo_y21_22) - min(employment_yo_y21_22))) %>% 
  # normalize rent to income
  mutate(n_avg_rent_to_income = (avg_rent_to_income- min(avg_rent_to_income)) / (max(avg_rent_to_income) - min(avg_rent_to_income))) %>% 
  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21))/(max(avg_occupancy_21) - min(avg_occupancy_21)))

@@

```{r}
# invert normalized rent to income ratio so lower values (more favorable) have higher scores
MSA_realEstate_data <- MSA_realEstate_data %>% 
  mutate(med_rent_to_income_inverted = 1 - n_med_rent_to_income) %>% 
  mutate(avg_rent_to_income_inverted = 1 - n_avg_rent_to_income)

@@

```{r}
# weights from owen
real_estate_weights <- c(0.35, 0.20, 0.15, 0.10, 0.10, 0.10)

#make real_estate_score column
MSA_realEstate_data <- MSA_realEstate_data %>% 
  mutate(real_estate_score = MSA) #this is just a place holder
```

```{r}
size <- dim(MSA_realEstate_data)

for (i in 1:size[1]) {
  MSA_x_data <- c(MSA_realEstate_data$n_pop_change[i],
                  MSA_realEstate_data$n_employ_change[i],
                  MSA_realEstate_data$n_avg_occupancy[i],
                  MSA_realEstate_data$n_rent_to_income[i],
                  MSA_realEstate_data$n_income_to_homeprice[i]
                  )
  MSA_realEstate_data$real_estate_score[i] = weighted.mean(MSA_x_data, real_estate_weights)
  }
```
